Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-23 paper/sections/proof-reconstruction/resolve.tex
[ paper ] Eq removed.

First, we define \rm{rassoc}$_{⬓}$ function in Eq.~(\ref{def:right-square})
to convert a disjunctions or a conjunction into its right-associative form.
We use $\rm{rassoc}_{∧}$ to convert conjunctions and $\rm{rassoc}_{∨}$ for
disjunctions.

@@

From now, we assume all propositions to be right-associative unless
otherwise stated.
The $\ffactor$ function in Eq.~(\ref{eq:factor-definition})
simplifies a special case of disjunction, the repeated disjuncts
(e.g., $\ffactor(φ ∨ φ) = φ$).
Notice that other cases like $φ ∨ (ψ ∨ φ)$ do no reduce to $ψ ∨ φ$.

@@

To construct a disjunction $ψ$ from a disjunction $φ$, we have used ideas
from the description in \citeauthor{bohme2010}~\cite{bohme2010} to prove
equality between nested disjunctions.
We define the $\rm{sbuild}_{∨}$ function in
Eq.~(\ref{eq:strong-build-or}).
that uses every disjunct from the source formula, $φ$, to build up the
target disjunction $ψ$.

>>>>

First, we define \rm{rassoc}$_{⬓}$ function in~(\ref{def:right-square})
to convert a disjunctions or a conjunction into its right-associative form.
We use $\rm{rassoc}_{∧}$ to convert conjunctions and $\rm{rassoc}_{∨}$ for
disjunctions.

@@

From now, we assume all propositions to be right-associative unless
otherwise stated.
The $\ffactor$ function in~(\ref{eq:factor-definition})
simplifies a special case of disjunction, the repeated disjuncts
(\eg, $\ffactor(φ ∨ φ) = φ$).
Notice that other cases like $φ ∨ (ψ ∨ φ)$ do no reduce to $ψ ∨ φ$.

@@

To construct a disjunction $ψ$ from a disjunction $φ$, we have used ideas
from the description in \citeauthor{bohme2010}~\cite{bohme2010} to prove
equality between nested disjunctions.
We define the $\rm{sbuild}_{∨}$ function in~(\ref{eq:strong-build-or}).
that uses every disjunct from the source formula, $φ$, to build up the
target disjunction $ψ$.

### 2017-10-23 paper/sections/proof-reconstruction/strip.tex
[ paper ] Eq removed.

To prove the correctness of the process mentioned above,
we need to establish some lemmas before introduce our main goal of this
part, Theorem~\ref{thm:thm-strip}.

@@

\begin{sketchproof}
Use induction on the cases defined by the outcome of the
$\fuh$ function in Eq.~(\ref{eq:uh-structured}).
\begin{itemize}
  \item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
  \item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
    \begin{equation*}
    \texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
    \end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
        argument.
\vskip 2mm
\begin{itemize}

@@

We now define the $\fstrip_0$ function in Eq.~\ref{eq:strip-unfixed} that
yields the conjunction of subgoals that implies the goal of the problem in
the \Metis \TSTP derivations.

\begin{equation}
\label{eq:strip-unfixed}
\begin{aligned}
&\hspace{.495mm}\fstrip_0 : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip_0 &(φ₁ ∧ φ₂)     &= \fuh~(\fstrip_0~φ₁) ∧ \fuh~(φ₁ ⇒ \fstrip_0~φ₂)\\
\fstrip_0 &(φ₁ ∨ φ₂)     &= \fuh~(¬ φ₁ ⇒ \fstrip_0~φ₂)\\
\fstrip_0 &(φ₁ ⇒ φ₂)     &= \fuh~(φ₁ ⇒ \fstrip_0~φ₂)\\
\fstrip_0 &(¬ (φ₁ ∧ φ₂)) &= \fuh~(φ₁ ⇒ \fstrip_0~¬ φ₂)\\
\fstrip_0 &(¬ (φ₁ ∨ φ₂)) &= \fuh~(\fstrip_0~¬ φ₁) ∧ \fuh~(¬ φ₁ ⇒ \fstrip_0~¬ φ₂)\\
\fstrip_0 &(¬ (φ₁ ⇒ φ₂)) &= \fuh~(\fstrip_0~φ₁) ∧ \fuh~(φ₁ ⇒ \fstrip_0~¬ φ₂)\\
\fstrip_0 &(¬ ¬ φ₁)      &= \fuh~(\fstrip_0~φ₁)\\
\fstrip_0 &(¬ ⊥)         &= ⊤\\
\fstrip_0 &(¬ ⊤)         &= ⊥\\
\fstrip_0 &φ             &= φ
\end{array}
\end{aligned}
\end{equation}

The function defined above has termination problems since it is not
a structural recursion. Therefore, we have reformulated it in
Eq.~\ref{eq:strip-fixed} in order to solve such an issue following
the technique described in Section~\ref{ssec:structural-recursion}.

@@

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $\text{n} : \Nat$ the complexity measure of the strip function in
Eq.~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fstrip~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula~$φ$ by following the cases in Eq.~(\ref{eq:strip-fixed}).
We present a straightforward case with double negation, a case with
conjunction connective, and last, the case with a negated disjunction.
We refer the reader to~\cite{AgdaMetis} for the complete proof in \Agda.

>>>>

To prove the correctness of the process mentioned above,
we need to establish some lemmas before introduce our main goal of this
part, Theorem~\ref{thm:strip}.

@@

\begin{sketchproof}
Use induction on the cases defined by the outcome of the
$\fuh$ function in~(\ref{eq:uh-structured}).
\begin{itemize}
  \item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
  \item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
    \begin{equation*}
    \texttt{∧⇒\rm{-to-}⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
    \end{equation*}
  \item For $n > 1$, we use induction on the structure of the second
        argument.
\vskip 2mm
\begin{itemize}

@@

We now define the $\fstripp$ function in~(\ref{eq:strip-unfixed}) that
yields the conjunction of subgoals that implies the goal of the problem in
the \Metis \TSTP derivations.

\begin{equation}
\label{eq:strip-unfixed}
\begin{aligned}
&\hspace{.495mm}\fstripp : \Prop → \Prop\\
&\begin{array}{lll}
\fstripp &(φ₁ ∧ φ₂)     &= \fuh~(\fstripp~φ₁) ∧ \fuh~(φ₁ ⇒ \fstripp~φ₂)\\
\fstripp &(φ₁ ∨ φ₂)     &= \fuh~(¬ φ₁ ⇒ \fstripp~φ₂)\\
\fstripp &(φ₁ ⇒ φ₂)     &= \fuh~(φ₁ ⇒ \fstripp~φ₂)\\
\fstripp &(¬ (φ₁ ∧ φ₂)) &= \fuh~(φ₁ ⇒ \fstripp~¬ φ₂)\\
\fstripp &(¬ (φ₁ ∨ φ₂)) &= \fuh~(\fstripp~¬ φ₁) ∧ \fuh~(¬ φ₁ ⇒ \fstripp~¬ φ₂)\\
\fstripp &(¬ (φ₁ ⇒ φ₂)) &= \fuh~(\fstripp~φ₁) ∧ \fuh~(φ₁ ⇒ \fstripp~¬ φ₂)\\
\fstripp &(¬ ¬ φ₁)      &= \fuh~(\fstripp~φ₁)\\
\fstripp &(¬ ⊥)         &= ⊤\\
\fstripp &(¬ ⊤)         &= ⊥\\
\fstripp &φ             &= φ
\end{array}
\end{aligned}
\end{equation}

The function defined above has termination problems since it is not
a structural recursion. Therefore, we have reformulated it
in~(\ref{eq:strip-fixed}) in order to solve such an issue following
the technique described in Section~\ref{ssec:structural-recursion}.

@@

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $\text{n} : \Nat$ the complexity measure of the strip function
in~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fstrip~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

\begin{sketchproof}
The proof is by induction on the structure of the
formula~$φ$ by following the cases in~(\ref{eq:strip-fixed}).
We present a straightforward case with double negation, a case with
conjunction connective, and last, the case with a negated disjunction.
We refer the reader to~\cite{AgdaMetis} for the complete proof in \Agda.
